---
title: "averageReuses.Rmd"
output: html_document
date: "2023-02-15"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(knitr)
```

## Calculate the average number of reuses of an item

Given a fixed probability of 'loss', i.e. not getting it back, what is the AVERAGE number of uses.

The average value of something is calculated as the 'expected value'. For a discrete distribution it is defined as Sum over all n's of (n*p(n)), where p(n) is the probability density of n (n is number of uses).  See here for more context: https://math.stackexchange.com/questions/1119872/on-average-how-many-times-must-i-roll-a-dice-until-i-get-a-6

For our case, p(n) is the probability that an item is lost on the nth use.
p(n) = (1 - p(not lost)^n) - (1 - p(not lost)^(n-1)) = p(not lost)^(n-1) - p(not lost)^n

The other approach is to notice that this problem is also a classical geometric distribution problem.  You'll get identical results following https://en.wikipedia.org/wiki/Geometric_distribution
The below median calculation comes from that.

```{r cars}
ploss <- 1/6
numsamples <- 1000
pnotlost <- 1-ploss

parr <- seq(0,numsamples,1)
#p1 <- (pnotlost^(parr-1) - pnotlost^parr)
p1 <- ploss*pnotlost^(parr-1)
plot(parr*p1, type = "l")

ExpectedValue <- sum(parr*p1)

cumsum <- c()
for (i in parr) {
  cumsum <- append(cumsum, sum(p1[1:i]))
}
cumsum <- cumsum/sum(p1)
medianval <- parr[max(which(cumsum<0.5))]
print(medianval)
nmed1 <- -1/log2(pnotlost) 
nmed2 <- log10(.5)/log10(pnotlost)
print(paste(nmed1,nmed2))
plot(cumsum, type ="l")
```

The expected value (mean or average) number of uses with a probability of loss of `r ploss` is roughly `r ExpectedValue`

For different values of loss probability, we can see that the average looks to be the inverse of the loss probability.
```{r lossfunc, results = 'asis'}
plossfunc <- function(ploss) {
  pnotlost <- 1-ploss

parr <- seq(1,numsamples,1)
#p1 <- pnotlost^(parr-1) - pnotlost^parr
p1 <- ploss*pnotlost^(parr-1)

ExpectedValue <- sum(p1*parr)
return (ExpectedValue)
}

probOfLoss <- seq(0.05, 0.5, 0.05)
averageUses <- c()
medianUses <- c()
for (x in probOfLoss) {
  averageUses <- append(averageUses,plossfunc(x))
  #medianUses <- append(medianUses, -1/log2(1-x) + 1)
  medianUses <- append(medianUses, log10(.5)/log10(1-x))
  #print(paste(x, plossfunc(x)))
}
df <- data.frame(probOfLoss, averageUses, medianUses)
kable(df)
```

